---
title: "Undirected Connectivity Measures"
author: "Zarrar Shehzad"
date: "07/11/2014"
output: html_document
---

I wanted to gather together various approaches to computing a full and partial correlation. I do that here with a brief comparison of the efficacy of the different approaches.


# Setup

Let's read in the data. Eventually I'd like to have this data at some central location so I could do all my business here below (e.g., having it up on a dropbox public folder). Anyway, for now this is all done locally on my computer. I will be using data from simulation 4 since it has the most nodes (50), which allows more comparisons to be made. For now, I will apply this to only one subject.

Let's read in the data. `net` is the ground truth network (nnodes x nnodes) and `ts` is the time-series data (ntpts x nnodes). I have a function `upper` to only use the upper half of a matrix to correspond to the use of the upper half in `net` to represent the connections.

```{r}
net   <- as.matrix(read.table("data/sim04/sub01_net.txt")) # reference
ts    <- as.matrix(read.table("data/sim04/sub01_ts.txt"))

# All the good stuff is in the upper half (i.e., actual connections)
upper <- function(x) x[upper.tri(x)]
```


# Connectivity

I will now compute both the full correlations and several different approaches for a partial/inverse covariance style matrix. The partial correlation approaches that I ended up choosing automatically estimate the lambda so that works well. I combine each of the correlation matrices together for easier analysis/comparison.

```{r}
# Function that calculates numerous correlation/covariance measures
compute_correlations <- function(ts) {
  cat("#full correlation\n")
  print(system.time(corr.full <- cor(ts)))
  
  cat("#inverse covariance\n")
  library(MASS)
  print(system.time(pcorr.icov <- -ginv(cov(ts))))
  
  cat("#inverse covariance with shrinkage (corpcor)\n")
  library(corpcor)
  print(system.time(pcorr.invcov <- -cov2cor(invcov.shrink(ts))))
  
  # ICOV: graphical lasso
  cat("#graphical lasso (glasso)\n")
  print(system.time(tmp <- glasso_cv(ts)))
  pcorr.glasso <- -cov2cor(tmp$wi)
  
  # ICOV: lasso
  cat("#partial correlations via adaptive lasso (parcor)\n")
  library(parcor)
  print(system.time(tmp <- adalasso.net(ts, k=5, verbose=T)))
  pcorr.lasso <- tmp$pcor.adalasso
  
  # ICOV: ridge
  cat("#partial correlations via ridge regression (parcor)\n")
  library(parcor)
  print(system.time(tmp <- ridge.net(ts, k=5, verbose=T)))
  pcorr.ridge <- tmp$pcor
  cat("\n")
  
  # ICOV: Ledoit-Wolf
  cat("#inverse covariance with ledoit-wolf shrinkage (BurStFin)\n")
  library(BurStFin)
  print(system.time(ctmp <- var.shrink.eqcor(ts)))
  itmp <- solve(ctmp)
  pcorr.lw <- -cov2cor(itmp)
  
  # Combine together into one list
  cmats <- list(
    full=corr.full, icov=pcorr.icov, invcov=pcorr.invcov, glasso=pcorr.glasso, lasso=pcorr.lasso, ridge=pcorr.ridge, lw=pcorr.lw
  )
}

cmats.orig <- compute_correlations(ts)
```


# Test Connectivity

I can now test how well these results do (without any thresholding) by getting the average positive and average negative results.

```{r}
ztrue <- function(x) mean(upper(x)[upper(net)!=0])
zfalse <- function(x) mean(upper(x)[upper(net)==0])
zmse <- function(x) mean(sqrt((upper(x)[upper(net)!=0] - upper(net)[upper(net)!=0])^2))

ret <- sapply(cmats.orig, function(cmat) {
  zt <- ztrue(cmat)
  zf <- zfalse(cmat)
  zm <- zmse(cmat)
  zd <- zt - zf
  c(true=zt, false=zf, diff=zd, mse=zm)
})

ret
```

```{r}
ztp <- function(x, thr=0) mean(abs(upper(x)>thr)[upper(net)!=0]) # true pos
zfp <- function(x, thr=0) mean(abs(upper(x)>thr)[upper(net)==0]) # fals pos

ret <- sapply(cmats.orig, function(cmat) {
  zt <- ztp(cmat)
  zf <- zfp(cmat)
  zd <- zt - zf
  c(true=zt, false=zf, diff=zd)
})

ret
```


```{r}
library(lassoscore)
ztp <- function(x) mean(x[upper(net)!=0]) # true pos
zfp <- function(x) mean(x[upper(net)==0]) # fals pos

tmp <- glasso_cv(ts)

# mbscore does well esp using fdr$qval
mbs <- mbscore(ts, tmp$rhomax)

table(me=upper(mbs$p.model)<0.001, ref=upper(net)!=0)
table(me=upper(mbs$p.sand)<0.01, ref=upper(net)!=0)

res <- fdrtool(upper(mbs$p.model), "pvalue", plot=F, verbose=F)$qval
c(true=ztp(res<0.05), false=zfp(res<0.05))
c(true=ztp(res<0.01), false=zfp(res<0.01))

res <- fdrtool(upper(mbs$p.model), "pvalue", plot=F, verbose=F)$lfdr
c(true=ztp(res<0.05), false=zfp(res<0.05))
c(true=ztp(res<0.01), false=zfp(res<0.01))

res <- fdrtool(upper(mbs$p.sand), "pvalue", plot=F, verbose=F)$qval
c(true=ztp(res<0.01), false=zfp(res<0.01))


# using glassoscore not as good
gls <- glassoscore(ts, tmp$rhomax)

table(me=upper(gls$p.model)<0.01, ref=upper(net)!=0)
table(me=upper(gls$p.sand)<0.01, ref=upper(net)!=0)

table(me=fdrtool(upper(gls$p.model), "pvalue", plot=F)$qval<0.05, ref=upper(net)!=0)
table(me=fdrtool(upper(gls$p.model), "pvalue", plot=F)$qval<0.01, ref=upper(net)!=0)
table(me=fdrtool(upper(gls$p.model), "pvalue", plot=F)$lfdr<0.05, ref=upper(net)!=0)
table(me=fdrtool(upper(gls$p.model), "pvalue", plot=F)$lfdr<0.01, ref=upper(net)!=0)
```
