---
title: "Connectivity via Random Data"
author: "Zarrar Shehzad"
date: "July 12, 2014"
output: html_document
---

Here, I want to test out different approaches for randomizing time-series data with the goal of creating a null model that can be used to set a threshold (e.g., p < 0.01) for correlations between time-series. I look specifically at thresholding on a different page. As an aside, if we are doing a task-based analysis, I wonder about the utility of using resting-state scans for the simulated data (e.g., randomly chop up in the same length as our trials). At the very least, this could serve as a good baseline even if it isn't very random.

Anyway, I will be using some of the Steve Smith et al (200?) resting-state simulations to see how well the different randomization techniques work in generating a threshold for my correlations. A good randomization technique should help set a threshold that minimizes false positives and/or false negatives while maximizing detection of true positives and/or true negatives.

The different techniques are:

* Block-Based Bootstrapping (tsboot)
* Block-Based Bootstrapping (tsbootstrap)
* Entropy-Based Bootstrapping (meboot)
* Wavelet-Based Bootstrapping (wmtsa)

In sum, you will notice that I choose the last three approaches above. The `tsboot` function overlaps with the `tsbootstrap` function and isn't as fast or easy to use.

## Setup

```{r}
# Functions needed for later
upper <- function(x) x[upper.tri(x)]
```


## Block-Based Resampling (boot) - NO

I am not totally what the `model` option in the `tsboot` function of the `boot` R package does but that is what will be used here. I think it has to do with resampling the data in time and with the `model` option, you don't have to choose the resample length.

```{r}
library(boot)

tmp <- tsboot(ts, function(x) upper(cor(x)), R=99, orig.t=T)
res <- rbind(tmp$t0, tmp$t)
mean(res[1,1] <= res[,1]) # significance (doesn't work here)

tmp <- tsboot(ts, function(x) upper(cor(x)), R=99, orig.t=T, sim="geom", l=20)
res <- rbind(tmp$t0, tmp$t)
mean(res[1,1] <= res[,1]) # significance

tmp <- tsboot(ts, function(x) upper(cor(x)), R=99, orig.t=T, sim="fixed", l=20)
res <- rbind(tmp$t0, tmp$t)
mean(res[1,1] <= res[,1]) # significance
```

I don't like that I need to select a block length here. I guess I would need to look into the literature to figure out what a good block length is. Well for now, I might put this one down.


## Block Based Resampling (tseries) - YES

The following is taken from the function help of `tsbootstrap` in the R package `tseries`: *"generates bootstrap samples for general stationary data and computes the bootstrap
estimate of standard error and bias if a statistic is given."*

This function is a little easier to understand than the `tsboot` one and they mention that it is 10-30 times faster than `tsboot` so this should be the function of choice. There are two ways the bootstrap replicates can be generated:

* **stationary**: stationary bootstrap scheme with mean block length set by user
* **block**: blockwise bootstrap with block length fixed by user

```{r}
tmp <- tsbootstrap(ts[,1], nb=1, type="stationary")
cor(ts[,1], ts[,2])
cor(tmp, ts[,2])

tmp <- tsbootstrap(ts[,1], nb=1, type="block")
cor(ts[,1], ts[,2])
cor(tmp, ts[,2])
```

So it seems one needs to bootstrap each time-series separately (which is fine) and I can then go about my business of doing my computations. In the above example, I didn't set the block length, which was auto-selected based on some criterian. Again, I would need to check the literature to see what an appropriate block length is or how to determine it or different block lengths that are required.


## Maximum Entropy Bootstrap Replicates (meboot) - YES

The details for the `meboot` function/package are pretty good so would suggest to look there.

```{r}
tmp <- meboot(ts[,1], reps=10, elaps=T)
cor(ts[,1], ts[,2])
cor(tmp$ensemble, ts[,2])
```

Seems easy to use but not totally sure like the other approaches what the strenghs are and what the weaknesses are here.


## Wavelet-Based Bootstrapping - Yes

Again I don't totally understand what this. From what I understand this does the bootstrap/randomization in frequency/wavelet domain. This keeps a lot of the properties of the time-series (not sure which ones though).

```{r}
list_to_mat <- function(l) sapply(l, function(v) v)
tmp <- wavBootstrap(ts[,1], n.realization=10)
cor(ts[,1], ts[,2])
cor(list_to_mat(tmp), ts[,2])
```


range for pierre stuff [2*ceil(sqrt(T)) 3*ceil(sqrt(T))]
code in python https://github.com/FCP-INDI/C-PAC/blob/c84bd3f5e8f1c6c5b19b2d171bb194cdd2f3e757/CPAC/basc/utils.py
code in r https://github.com/grayclhn/oosanalysis-R-library/blob/9950a4a38a842b936d6a58d70c81c66e6579469d/R/bootindex.R
paper http://www.sciencedirect.com/science/article/pii/S1053811910002697
review http://books.google.com/books?hl=en&lr=&id=NOCYBtPzQFIC&oi=fnd&pg=PA3&dq=bootstrap+timeseries&ots=eXJxyqoNh-&sig=d4FbmPWkQTSHc2baNAz63zWxHv4

opt.boot.lens <- np::b.star(ts, round=T)[,2]
unique(opt.boot.lens[opt.boot.lens>1])


```{r}
zboots.circularblock <- function(ts, reps=1, drop=FALSE) {
  suppressPackageStartupMessages(library(niftir)) # for progressbar
  resamp.ts  <- array(NA, c(dim(ts), reps))
  pb <- progressbar(reps)
  for (i in 1:reps) {
    if (!(i %% 10)) update(pb, i)
    block.lens <- zboots.circularblock.lens(ts) # block lengths
    block.inds <- zboots.circularblock.inds(nrow(ts), block.lens) # block indices
    resamp.ts[,,i]  <- ts[block.inds,]
  }
  end(pb)
  if ((reps == 1) && (drop == T)) {
    return(resamp.ts[,,1])
  } else {
    return(resamp.ts)
  }
}

# block lengths
zboots.circularblock.lens <- function(ts) {
  # First get optimal length via Patton et al. (2009)
  # but keep minimal length to >1 per Bellec et al. (?)
  library(np)
  nbs1  <- b.star(ts, round=T)
  nbs1  <- sort(unique(nbs1[,2]))
  nbs1  <- nbs1[nbs1>1]
  
  # Second we get recommended lengths from Bellec et al. (?)
  nobs  <- nrow(ts)
  nbs2  <- c(2,3) * ceiling(sqrt(nobs))
  
  # Combine those together
  blocklens <- sort(unique(c(nbs1, nbs2)))
  
  return(blocklens)
}

# borrowed/updated from
# https://github.com/grayclhn/oosanalysis-R-library/blob/9950a4a38a842b936d6a58d70c81c66e6579469d/R/bootindex.R
zboots.circularblock.inds <- function(nobs, blocklens) {
  # randomly select one of the block lengths if more than 1
  blocklen <- ifelse(length(blocklens)>1, sample(blocklens, 1), blocklens)
  
  blockstarts <- sample(seq_len(nobs), ceiling(nobs / blocklen),
                        replace = TRUE)
  indices <-
    as.vector(sapply(blockstarts, function(s)
                     (s + seq_len(blocklength) - 2) %% nobs + 1))[seq_len(nobs)]
  attr(indices, "starts") <- blockstarts
  attr(indices, "length") <- blocklen
  indices
}
```